\section{Conclusion}

The goal of this thesis has been to present and analyze different matching mechanisms and their properties for one-to-many matching scenarios with one-sided preferences. A variety of optimality criteria exist, but in the end, it is evident that, none of the algorithms presented produce matchings that fulfill all of the optimality criteria. 

The experiments have shown that the resulting matchings vary quite a lot by algorithm and structure of the instance. In summary, the Hungarian algorithm usually outperformed all other algorithms on most metrics; but, it also had by far the highest runtime. To improve the runtime, different algorithms can be used, such as Rank-Max, that also compute profile-based matchings in almost linear time. Another insight from the experiments was that the Popular-CHA algorithm failed to find matchings for a majority of the instances, however we discussed a simple modified version of the algorithm that usually produced better results than RSD and Max-PaCHA, while having a much lower runtime than the Hungarian algorithm. Still, according to the results from the experiment, \emph{Popularity} seemed to be a less desirable criteria, due to the fact that popular matchings often only optimized the match for a subset of the students, while leaving the other students with a bad or non-existent match. Therefore, we can recommend the Hungarian or Rank-Max algorithm for the use case of student-seminar matching. To improve on performance and lower the bar for strategic manipulation, a matching system could require students to supply at least $k$ preferences. This would also partially eliminate the Hungarian's algorithm biggest disadvantage, which was that students are encouraged to provide short preference lists.

As part of this thesis, we have also presented an interactive web-system that allows university administrations to manage and solve such matching problems. The system relies on the C++ algorithms that were used for the benchmark and should offer good performance for typical instances.  